Selective stock enhancement undermines community stability

### 2022-01-14 rmd/manuscript.Rmd
edit: manuscript

@@

Anthropogenic influences have caused global biodiversity crisis, such that human interventions are critical to the sustainable management of natural resources. Captive breeding is a common practice to conserve or enhance wild populations of diverse plant and animal taxa. Although there is growing recognition that stock enhancement with captive-bred organisms entails ecological risks (e.g., the accumulation of deleterious alleles), this method is still pervasive in conservation and fisheries. In fisheries for example, billions of hatchery-reared individuals are produced and released annually across the globe. The widespread use of captive breeding is perhaps because of the "myth" that numerical benefits of released individuals may exceed the associated risks.

Current debates, however, overlook the fact that we have rarely assessed the community-wide impact of stock enhancement. Species are all embedded in the complex web of interacting organisms, and stable coexistence of competing species through density-dependent feedback underpins emergent stability of ecological communities. Stock enhancement may disrupt the sensitive balance of species interactions because it introduces unnaturally high numbers of individuals into the wild. Hence, this form of species management may have significant consequences for community dynamics. Evidence for this hypothesis is lacking, however.

Here, we use a simple theory to show that stock enhancement undermines long-term community stability, which we define here as the relative size of fluctuations in total community density over time. Our theory illuminates that stock enhancement interferes the internal regulation process of self-sustainable communities, compromising the stabilizing mechanism emerging from species niche difference. The present study further demonstrates the relevance of this general theory to natural systems by illustrating its congruence with Japanese stream fish communities, where \>10 millions of hatchery masu salmon (*Oncorhynchus masou masou*) are released annually for fisheries, recreation, and conservation purposes. Our findings paint a bleak picture for stock enhancement, reinforcing the recurrent calls for alternative management strategies.

We used a multispecies Ricker model to simulate community dynamics with selective stock enhancement of a constituent species (species 1). Specifically, population density of species $i$ at time $t+1$, $N_{i,t+1}$, is modeled as:

@@

where $r_i$ is the intrinsic growth rate, $\alpha_{ij}$ the competition coefficient of species $j$ on species $i$ (interspecific competition drawn from an exponential distribution as $\alpha_{ij,i \ne j} \sim Exp(1/\bar{\alpha})$ whereas $\alpha_{ii}=1$), $K$ the community-level carrying capacity in the absence of niche difference, $R_t$ the number of released individuals for the enhancement of species 1 (set constant as $R_t = R$ for simplicity), and $\epsilon_{i,t}$ the species responses to stochastic environmental fluctuations that follow an normal distribution with a mean of zero and a standard deviation of $\sigma_{\epsilon}$. The parameter $\phi_i$ takes the following values:

@@

$f_R~(\ge 0)$ is the density-independent survival of released individuals relative to wild individuals (the relative fitness). Therefore, the model accounts for fitness difference of captive-bred individuals due to genetic effects and/or phenotypic plasticity when considering the reproductive contribution of released individuals to the next generation of the enhanced species.

In our simulation, we used `r nrow(param_set)` simulation scenarios with varied values of the following ecological parameters: mean strength of interspecific competition $\bar{\alpha}$, intrinsic growth of the enhanced species $r_1$, relative fitness of captive-bred individuals $f_R$, and degree of environmental stochasticity $\sigma_{\epsilon}$. Under each scenario (i.e., parameter combination), we ran 1400 time steps of 1000 independent communities (i.e., simulation replicates) with varying $R$. The last 1000 time steps were used to obtain the coefficient of variation, temporal mean, and temporal standard deviation of total community density ($\sum_i^SN_i$). Temporal means and standard deviations were calculated separately for the enhanced species $N_1$ and the unenhanced species $\sum_{i,i \ne 1}^S N_i$ to infer mechanisms that underlie community-level responses.

Our model predicted a destabilizing effect of selective stock enhancement on ecological communities as illustrated by increased CV with increasing numbers of released individuals $R$ (Figure XX). This pattern stemmed mainly from the reduced means of enhanced and unenhanced species groups (Figure XX) that are driven by different mechanisms. For enhanced species, released individuals induced overcompensation, where the negative effect of reduced per-capita reproduction due to competition exceeds the numerical benefit of released individuals. Therefore, the enhanced species reached the lower long-term average of population density with higher numbers of released individuals. Meanwhile, the reduced mean of unenhanced species is caused by interspecific competition between enhanced and unenhanced species. Combined, the long-term average of the total community density decreased more sharply than individual species groups with stock enhancement (XX). The SDs showed a similar trend, but the relationship was flatter at the community level. Since a CV is a ratio of an SD to a mean, the steeper relationship of the mean community density has led to the increased CV. These patterns were qualitatively similar across ecological scenarios, even when the interspecific competition is very weak.

The destabilizing effect emerges because stock enhancement undermines the balance of species interactions that underpin community stability. Stable coexistence of competing species rests largely on niche difference in consumable resource use ($\alpha_{ij, i \ne j} < 1$) because it creates a situation where numerically dominant species undergoes stronger intraspecific competition. However, stock enhancement is externally controlled and the number of release is not subject to density-dependent regulation. Such excessive individuals impose additional intra- and interspecific competition. Consequently, stock enhancement compromises the ecological process giving rise to "overyielding", in which total community density of an ecologically diverse community exceeds what would be expected in a single species community ($\sum_i^SN_i > K$). In an extreme scenario, intensive stock enhancement extirpated competing species, reducing biodiversity in the recipient community.

We must note that stock enhancement helps increase the density of the enhanced species when the intrinsic population growth rate is significantly small ($r_1 < 0.5$). This result is reasonable because some of the best evidence for successful stock enhancement comes from critically endangered species. Yet, the community-level effect is still destabilizing due to the intensified competition with unenhanced species. Therefore, managers should consider the propagating effect of captive breeding when implementing their programs.

@@

To demonstrate the relevance of our general theory to natural systems, we assessed the potential impacts of selective stock enhancement of masu salmon on long-term stability of stream fish communities. Masu salmon is one of the most valuable freshwater resources in Hokkaido Island, Japan, and XX leads extensive hatchery programs across the island. In protected watersheds where any form of exploitation is prohibited (including recreational angling), long-term monitoring has been conducted with consistent sampling methods over the past two decades along with official records of stock enhancement. Therefore, the dataset sets the stage for a "natural experiment" to test our theoretical predictions. In this study, we used time-series data of stream fish communities at XX sites within XX protected watersheds (see Methods for selection criterion). Using hierarchical Bayesian models, we quantified the relationship between community dynamics and the number of fish released while accounting for potential confounding factors (climates and local abiotic factors).

As our theory predicts, stream fish communities showed greater temporal fluctuations (higher CV) in watersheds with greater numbers of fish released. The positive relationship between the CV and the number of fish released was strongly supported by a hierarchical linear model, in which the probability of the regression coefficient being positive was XXX. This result was associated with the reduced long-term averages of population densities of enhanced (masu salmon) and unenhanced fish groups. Both fish groups showed low population densities with high numbers of fish released (STATS), although statistical support for this relationship in the individual fish groups was weak to moderate (STATS). In the meantime, the SDs of population densities had vague relationships with the number of fish released, and neither positive or negative relationship was supported by statistical analysis.

Our study system satisfies key theoretical assumptions that are required for the destabilizing effect of stock enhancement to emerge. The first assumption is that released fish compete with wild fish. This type of competition is very likely to occur in the study streams. Masu salmon typically spend one year in freshwater before migrating to the ocean, and a wealth of evidence suggests that this species compete with other species of stream fish. In particular, masu salmon raised in hatchery environments are more aggressive than wild individuals, increasing the likelihood of intense intra- and interspecific competition in the wild. Second, our theory predicts the emergence of destabilizing effect when the number of fish released is externally controlled. This is exactly what was observed in our study system. In the protected watersheds, there was no apparent congruence between the number of fish released and the population density of masu salmon (XX). Instead, these released fish are likely controlled by socioeconomic factors, such as budget allocation to hatchery programs in this region. As such, released fish are probably "excessive" given the density-dependent regulation of natural communities, imposing intense resource competition that would otherwise not exist.

Like any non-experimental studies, we cannot exclude the possibility that our results were driven by spurious correlations. However, this is very unlikely because we have controlled potential confounding factors, including climatic and land use variables. In addition, our results were in close agreement with theoretical predictions, which are free from any confounding factors. Hence, our empirical results are sufficiently reliable.

@@

Although captive breeding programs pay close attention to the fate of captive-bred individuals in the wild, it is rare to consider the self-regulation process of biodiversity. The results of this study demonstrate that the ignorance of this critical process will erode the long-term persistence of the recipient community. The destabilizing effect of stock enhancement is likely to be pervasive across terrestrial and aquatic taxa because negative density dependence is a general regulating mechanism of biodiversity. We anticipate that the undesired influence of stock enhancement may be even more apparent in degraded habitats due to the compromised capacity to support biodiversity. Therefore, it is essential for us to shift our focus to the preservation and restoration efforts of natural habitats. For example, a modern conservation initiative underscores the importance of conserving a diverse habitat portfolio that may buffer the impacts of rapid environmental changes. Without a broader framework that appreciates the ecological integrity of self-sustaining communities, stock enhancement will never be effective but impairs biodiversity.

@@

We developed a multispecies Ricker model to simulate community dynamics with selective stock enhancement of a particular species. Population density of species $i$ at time $t+1$, $N_{i,t+1}$, is modeled as:

$$
N_{i,t+1} = [N_{i,t}+\phi_i R_t]exp[r_i(1-\frac{N_{i,t}+\alpha_{i1} R_t + \sum_{j=1}^S \alpha_{ij}N_{j,t}}{K_i})]exp(\epsilon_{i,t})
$$

where $r_i$ is the intrinsic growth rate, $\alpha_{ij}$ the competition coefficient between species $i$ and $j$ (intraspecific competition if $i=j$; $\alpha_{ii}=1$), $K_i$ the carrying capacity, $R_t$ the number of released individuals for the stock enhancement of species 1, and $\epsilon_{i,t}$ the species responses to stochastic environmental fluctuations. The parameter $\phi_i$ takes a value of one when $i=1$ (otherwise zero) to represent the reproductive contribution of released individuals to the next generation of the enhanced species. The species-specific intrinsic growth rates and competition coefficients were drawn randomly from a uniform and exponential distribution, respectively, to allow for interspecific variation as $r_i \sim Unif(0.5, r_{max})$ and $\alpha_{ij, i \ne j} \sim Exp(\mu_{\alpha})$. The number of released individuals was assumed to be constant across time steps ($R_t = R$). Environmental stochasticity was drawn from a normal distribution as $\epsilon_{i,t} \sim Normal(\mu_{\epsilon}, \sigma_{\epsilon}^2)$

We varied values of four parameters: $S =$ `r unique(param_set$n_species)`; $r_{max}=$ `r unique(param_set$r_max)`; $\mu_{\alpha} =$ `r unique(param_set$alpha)`; $\sigma_\epsilon =$ `r unique(param_set$sd_env)`. For simplicity, we assumed identical carrying capacities among simulated species as $K_i =$ `r unique(param_set$k)`. This setup resulted in `r nrow(param_set)` sets of parameter combinations that cover a range of ecological scenarios.

Under each parameter combination, we ran 1400 time steps of 1000 independent communities (i.e., simulation replicates) with the gradient of the number of released individuals $R$ (0 - 500). We initialized the community with populations of each species drawn from a poisson distribution with a mean of five. We repeated the seeding procedure every 10 time steps over the first 200 time steps to allow species to establish populations and reach an equilibrium without (initialization period). After the initialization period, we released $R$ individuals of the enhanced species every time step over the next 200 time steps to reach a new equilibrium with selective stock enhancement (burn-in period). We saved the last 1000 time steps to obtain the relative temporal fluctuations (i.e., the coefficient of variation) of the whole community ($\sum_i^SN_j$), the enhanced species ($N_1$), and the unenhanced species ($\sum_{i,i \ne 1}^S N_i$). We also calculated temporal means and standard deviations for each species group. Values of simulation parameters were summarized in XXX. We performed simulations using R version XXX.

@@

**Time-series data.** We assembled time-series data of fish abundance across XX sites within XX protected watersheds of Hokkaido Island, Japan. In these watersheds, aquatic species are protected and no harvest is allowed. At these sites, the long-term monitoring program began in 1971 but effective sampling methods (a combination of electrofishing and cast net sampling) were implemented in 1999. Data were collected in summer to fall with irregular sampling intervals (0 to 3 year intervals for most cases). After careful data selection (Supporting Information), we used time-series data of fish abundance at XX sites during a period of 1999-2019. ADD SOME DETAILS

**Fish stocking.** At the protected watersheds, masu salmon were stocked to enhance the wild populations since XXX. A mixture of captive-bred and wild-born individuals were released in spring (fry [age 0+] and smolt [age 1+] stages) and fall (juvenile stage [age 0+]). We examined historical records of fish stocking and obtained annual stocking data (the number of individuals released) for a period of 1999-2019 at each watershed. We were unable to extract stocking data at the site level.

**Environmental data.** At each sampling site,we collected the following data as potential covariates: upstream watershed area (km^2^; a proxy for stream size), proportional land use in the upstream watershed (forest, agriculture, urban), local species richness, local climates (annual mean air temperature [unit] and cumulative precipitation [mm]), and ocean productivity (Chlorophyll a [unit]).

@@

Our goal is to compare temporal dynamics of fish communities across sites. However, the data are not directly comparable because of artificial observation errors (e.g., different observers) and missing observations. To confront this challenge, we developed Bayesian state space models for three species groups separately: (1) whole community, summed abundance of all species, (2) enhanced species, abundance of masu salmon, (3) unenhanced species, summed abundance of all species except masu salmon. Bayesian state space models is best suited for our analysis because it can account for observation errors while imputing missing values given the long-term trend at the site. The model is composed of observation and state models as described below.

@@

$n_{s,t}$ is the "true" fish density excluding fish released in the spring, $Fry_{w(s),t}$ the number of salmon fry released in spring in watershed $w$ within which site $s$ is located, and $\beta_s$ the site-specific effect of released salmon fry on the observed fish density. The parameter $e^{\epsilon_{s,t}^{obs}}$ is a random error variable drawn from a log-normal distribution as $\epsilon_{s,t}^{obs} \sim Normal(0, \sigma_{obs,s}^2)$. The inclusion of this random variable allows us to account for site-specific and year-specific observation errors. The parameter $\psi$ equals one when we model the unenhanced species group (otherwise $\psi = 1$). We included $\psi$ to exclude the term $\beta~Fry_{w(j),t}$ because the unenhanced species group does not include masu salmon.

@@

$$
ln~n_{s,t+1} = ln~r_{s} + ln~n_{s,t} + \epsilon_{s,t}^{state}
$$where $ln~r_{s}$ is the average rate of change at site $s$ and $\epsilon_{s,t}^{state}$ is the process error that is assumed to follow a normal distribution as $\epsilon_{s,t}^{state} \sim Normal(0, \sigma_{state,s}^2)$. The site-specific rate of change is random draws from a normal distribution as $ln~r_s \sim Normal(\mu_r, \sigma_r^2)$, assuming that community dynamics across Hokkaido have a shared temporal trend to some degree (the degree of shared trend is controlled by the standard deviation $\sigma_r$). This hierarchical structure allows for improved estimates of $ln~r_s$ by sharing information across sites. We used median estimates of fish density $n_{s,t}$ to calculate temporal CV, mean ($\mu$), and SD ($\sigma$).

>>>>

n_sim <- df_theory %>% 
  mutate(n_sim = n_warmup + n_burnin + n_timestep) %>% 
  pull(n_sim) %>% 
  unique()

@@

Anthropogenic influences have caused global biodiversity crisis, such that human interventions are critical to the sustainable management of natural resources [@cardinaleBiodiversityLossIts2012]. Captive breeding is a common practice to conserve or enhance wild populations of diverse plant and animal taxa [@snyderLimitationsCaptiveBreeding1996]. Although there is growing recognition that stock enhancement with captive-bred organisms entails ecological risks (e.g., the accumulation of deleterious alleles) [@arakiGeneticEffectsCaptive2007a; @snyderLimitationsCaptiveBreeding1996; @griffithsCaptiveBreedingReintroduction2008], this method is still pervasive in conservation and fisheries. In fisheries for example, billions of hatchery-reared individuals are produced and released annually across the globe [@kitadaEconomicEcologicalGenetic2018]. The widespread use of captive breeding is perhaps because of the "myth" that numerical benefits of released individuals may exceed the associated risks.

Current debates, however, overlook the fact that we have rarely assessed the community-wide impact of stock enhancement. Species are all embedded in the complex web of interacting organisms, and stable coexistence of competing species through density-dependent feedback underpins emergent stability of ecological communities [@tilmanEcologicalConsequencesChanges1999a]. Stock enhancement may disrupt the sensitive balance of species interactions because it introduces unnaturally high numbers of individuals into the wild. Hence, this form of species management may have significant consequences for community dynamics. Evidence for this hypothesis is lacking, however.

Here, we use a simple theory to show that stock enhancement undermines long-term community stability, which we define here as the relative size of fluctuations in total community density over time [@tilmanEcologicalConsequencesChanges1999a]. Our theory illuminates that stock enhancement interferes with the internal regulation process of self-sustainable communities, compromising the stabilizing mechanism emerging from species niche differences. The present study further demonstrates the relevance of this general theory to natural systems by illustrating its congruence with Japanese stream fish communities, where \>10 million hatchery masu salmon (*Oncorhynchus masou masou*) are released annually for fisheries, recreation, and conservation purposes. Our findings paint a bleak picture for stock enhancement, reinforcing the recurrent calls for alternative management strategies.

## Theoretical predictions

We used a multispecies Ricker model [@fowlerSpeciesDynamicsAlter2012] to simulate community dynamics with selective stock enhancement of a constituent species (species 1). Specifically, population density of species $i$ at time $t+1$, $N_{i,t+1}$, is modeled as:

@@

where $r_i$ is the intrinsic growth rate, $\alpha_{ij}$ the competition coefficient of species $j$ on species $i$ (interspecific competition drawn from an exponential distribution as $\alpha_{ij,i \ne j} \sim Exp(1/\bar{\alpha})$ whereas intraspecific competition $\alpha_{ii}=1$), $K$ the community-level carrying capacity in the absence of niche difference, $R_t$ the number of released individuals for the enhancement of species 1 (set constant as $R_t = R$ for simplicity), and $\epsilon_{i,t}$ the species responses to stochastic environmental fluctuations that follow an normal distribution with a mean of zero and a standard deviation of $\sigma_{\epsilon}$. The parameter $\phi_i$ controls the relative fitness of captive-bred individuals as follows:

@@

$f_R~(\ge 0)$ is the density-independent survival of captive-bred individuals relative to wild individuals. Therefore, the model accounts for fitness difference of captive-bred individuals due to genetic effects and/or phenotypic plasticity when considering the reproductive contribution of released individuals to the next generation of the enhanced species.

In our simulation, we used `r nrow(param_set)` simulation scenarios with varied values of the following ecological parameters: mean strength of interspecific competition $\bar{\alpha}$, intrinsic growth of the enhanced species $r_1$, relative fitness of captive-bred individuals $f_R$, and degree of environmental stochasticity $\sigma_{\epsilon}$. Under each scenario with a unique parameter combination, we ran `r n_sim` time steps of `r max(df_theory$n_rep)` independent communities (i.e., simulation replicates) with varying numbers of released individuals $R$. The last `r unique(df_theory$n_timestep)` time steps were used to obtain the coefficient of variation, temporal mean, and temporal standard deviation of total community density ($\sum_i^SN_i$). Temporal means and standard deviations were calculated separately for the enhanced species $N_1$ and the unenhanced species $\sum_{i,i \ne 1}^S N_i$ to infer mechanisms that underlie community-level responses.

Our model predicted a destabilizing effect of selective stock enhancement on ecological communities as illustrated by increased CV with increasing numbers of released individuals $R$ (Figure XX). This pattern stemmed mainly from the reduced means of enhanced and unenhanced species groups (Figure XX) that are driven by different mechanisms. For enhanced species, released individuals induced overcompensation, where the negative effect of reduced per-capita reproduction due to competition exceeds the numerical benefit of released individuals [@satakeStockingCaptivebredFish2012]. Therefore, the enhanced species reached the lower long-term average of population density with higher numbers of released individuals. Meanwhile, interspecific competition between enhanced and unenhanced species caused the reduced mean of unenhanced species, leading to decreased numbers of persisting species at high levels of stock enhancement. Combined, the long-term average of the total community density decreased more sharply than individual species groups with stock enhancement (XX). The SDs showed a similar trend, but the relationship was flatter at the community level. Since a CV is a ratio of an SD to a mean, the steeper relationship of the mean community density has led to the increased CV. These patterns were qualitatively similar across ecological scenarios, even when the interspecific competition is very weak.

The destabilizing effect emerges because stock enhancement undermines the balance of species interactions that underpin community stability. Stable coexistence of competing species rests largely on niche difference in consumable resource use ($\alpha_{ij, i \ne j} < 1$) because it creates a situation where numerically dominant species undergoes stronger intraspecific competition [@lettenLinkingModernCoexistence2017]. However, stock enhancement is externally controlled and the number of release is not subject to density-dependent regulation. Such excessive individuals impose additional intra- and interspecific competition. Consequently, stock enhancement compromised the ecological process giving rise to "overyielding" [@tilmanEcologicalConsequencesChanges1999a], in which total community density of an ecologically diverse community exceeds what would be expected in a single species community ($\sum_i^SN_i > K$).

We must note that stock enhancement helps increase the density of the enhanced species when the intrinsic population growth rate is significantly small ($r_1 = 0.5$). This result is reasonable because some of the best evidence for successful stock enhancement comes from critically endangered species [@johnsonGeneticRestorationFlorida2010; @oppelPopulationReinforcementDemographic2021]. Yet, the community-level effect is still destabilizing due to the intensified competition with unenhanced species. It is therefore critical to consider the propagating effect when implementing captive breeding programs.

@@

To demonstrate the relevance of our general theory to natural systems, we assessed the potential impacts of selective stock enhancement of masu salmon on the long-term stability of stream fish communities. Masu salmon is one of the most valuable freshwater resources in Hokkaido Island, Japan, and XX leads extensive hatchery programs across the island. In protected watersheds where any form of exploitation is prohibited (including recreational angling), long-term monitoring has been conducted with consistent sampling methods over the past two decades along with official records of stock enhancement [@teruiMetapopulationStabilityBranching2018]. Therefore, the study system sets the stage for a "natural experiment" to test our theoretical predictions. In this study, we used time-series data of stream fish communities at XX sites within XX protected watersheds (see Methods for selection criterion). Using hierarchical Bayesian models, we quantified the relationship between community dynamics and the number of fish released while accounting for potential confounding effects of climates and local abiotic factors.

As our theory predicts, stream fish communities showed greater temporal fluctuations (higher CV) in watersheds with greater numbers of fish released. The positive relationship between the CV and the number of fish released was strongly supported by a hierarchical linear model, in which the probability of the regression coefficient being positive was XXX. This pattern was associated with the reduced long-term averages of population densities of enhanced (masu salmon) and unenhanced fish groups, although statistical support for this relationship was weak to moderate (STATS). In the meantime, the SDs of population densities had vague relationships with the number of fish released, and neither positive or negative relationship was supported by statistical analysis. These results are in good agreement with previous studies. Negative or vague population response to stock enhancement has been widely observed in other fish and invertebrate species, implying limited effectiveness of this method as a management tool. Our analysis of long-term community data further illuminated the potential negative impact of stock enhancement on unenhanced species and expanded our understanding to community-level response to stock enhancement.

In our theory, the observed pattern emerges when negative density dependence operates in the recipient community, and this assumption is well supported by previous observations. Masu salmon typically spend one year in freshwater before migrating to the ocean, and a wealth of evidence suggests that released masu salmon compete with wild individuals of masu salmon [@sahashiEvaluationContributionHatchery2015a] and other stream fishes [@hasegawaEffectsHatcheryChum2014; @hasegawaWildMasuSalmon2018]. In particular, masu salmon raised in hatchery environments are larger and more aggressive than wild individuals, increasing the likelihood of intense intra- and interspecific competition in the wild [@hasegawaWildMasuSalmon2018]. Importantly, the number of fish released is likely controlled by socioeconomic factors (e.g., budget allocation) because there was no apparent congruence between the number of fish released and the population density of masu salmon (XX). As such, released fish are "excessive" given the density-dependent regulation of recipient communities and may impose resource competition that would otherwise not exist.

Reduced fitness of captive-bred individuals may also play a role in producing the observed community response. As with other salmonid species, ocean-migrating adults of captive-bred masu salmon show lower return rates to the spawning river and therefore less contribute to the next generation (miyakoshi?). These processes, i.e., negative density dependence and fitness disadvantage of captive-bred fish, may act in concert to reduce the long-term average of community density, undermining community stability in the long term.

Like any non-experimental studies, we cannot exclude the possibility that our results were driven by spurious correlations. However, in the protected watersheds, any exploitation of aquatic species and alteration of in-stream habitats (e.g., construction of river-crossing structure) are strictly regulated; therefore, it is difficult to envision that unmeasured human influences played a major role in the observed relationships. Further, we have statistically controlled natural variation in environmental factors, including climatic and land use variables. Hence, we are confident in our statistical inference, and this unique study setup may have helped produce the close agreement between theory and empirical patterns.

@@

Although captive breeding programs pay close attention to the fate of captive-bred individuals in the wild, it is rare to consider the self-regulation process of biodiversity. The results of this study demonstrate that the ignorance of this critical process will erode the long-term persistence of the recipient community. The destabilizing effect of stock enhancement is likely to be pervasive across terrestrial and aquatic taxa because negative density dependence is a general regulating mechanism of ecological communities.

We anticipate that the undesired influence of stock enhancement may be more apparent in degraded habitats as the compromised capacity to support biodiversity may impose greater competition among constituent species. Therefore, we should shift our management efforts to the preservation and restoration of natural habitats with a broader scope to conserve biodiversity. For example, a modern conservation initiative underscores the importance of conserving a diverse habitat portfolio that may buffer the impacts of rapid environmental changes. Without a comprehensive framework that appreciates the ecological integrity of self-sustaining communities, stock enhancement will never be effective but impairs biodiversity.

@@

We developed a multispecies Ricker model to simulate community dynamics with selective stock enhancement of a constituent species (species 1). Population density of species $i$ at time $t+1$, $N_{i,t+1}$, is modeled as:

$$
N_{i,t+1} = [N_{i,t}+\phi_i R_t]exp[r_i(1-\frac{N_{i,t}+\alpha_{i1} R_t + \sum_{j=1}^S \alpha_{ij}N_{j,t}}{K})]exp(\epsilon_{i,t})
$$

where $r_i$ is the intrinsic growth rate, $\alpha_{ij}$ the competition coefficient between species $i$ and $j$, $K$ the community-level carrying capacity in the absence of niche difference, $R_t$ the number of released individuals for the stock enhancement of species 1 (set constant as $R_t = R$), and $\epsilon_{i,t}$ the species responses to stochastic environmental fluctuations that obey an normal distribution with a mean of zero and a standard deviation of $\sigma_{\epsilon}$. The parameter $\phi_i$ controls the relative fitness of captive-bred individuals as follows:

$f_R~(\ge 0)$ is the density-independent survival of captive-bred individuals relative to wild individuals.

We varied values of the following ecological parameters: mean strength of interspecific competition ($\bar{\alpha}=$ `r unique(param_set$alpha)`), intrinsic growth of the enhanced species ($r_1 =$ `r unique(param_set$r1)`), relative fitness of captive-bred individuals ($f_R=$ `r unique(param_set$phi)`), and degree of environmental stochasticity ($\sigma_{\epsilon}=$ `r unique(param_set$sd_env)`). Other parameters were set constant or drawn randomly from a probability distribution: $S=$ `r unique(param_set$n_species)` , $K=$ `r unique(param_set$k)` , $r_i =$ $Unif$(`r unique(param_set$r_min)` , `r unique(param_set$r_max)`) ($i \ne 1$), $\alpha_{ii} = 1$, $\alpha_{ij,i \ne j} \sim Exp(\bar{\alpha})$. This simulation setup resulted in `r nrow(param_set)` sets of parameter combinations that cover a range of ecological scenarios.

Under each parameter combination, we ran `r n_sim` time steps of `r max(df_theory$n_rep)` independent communities (i.e., simulation replicates). The number of released individuals $R$ was drawn randomly from a uniform distribution for each simulation replicate as $R \sim Unif(0,500)$. We initialized the community with populations of each species drawn from a poisson distribution with a mean of five. We repeated the seeding procedure every 10 time steps over the first `r unique(df_theory$n_warmup)` time steps to allow species to establish populations and reach an equilibrium without stock enhancement (initialization period). After the initialization period, we released $R$ individuals of the enhanced species every time step over the next `r unique(df_theory$n_burnin)` time steps to reach a new equilibrium with selective stock enhancement (burn-in period). After the burn-in period, we continued the simulation run with stock enhancement and saved the last `r unique(df_theory$n_timestep)` time steps to obtain the relative temporal fluctuations (i.e., the coefficient of variation) of the whole community ($\sum_i^SN_j$), the enhanced species ($N_1$), and the unenhanced species ($\sum_{i,i \ne 1}^S N_i$). We also calculated temporal means and standard deviations for each species group. Values of simulation parameters were summarized in XXX. We performed simulations using R version XXX.

@@

**Time-series data.** We assembled time-series data of fish abundance across XX sites within XX protected watersheds of Hokkaido Island, Japan. In these watersheds, aquatic species are protected and no harvest is allowed. At these sites, the long-term monitoring program began in 1971 but effective sampling methods (a combination of electrofishing and cast net sampling) were implemented in 1999. Data were collected in summer to fall with irregular sampling intervals (0 to 3 year intervals for most cases). After careful data selection (Supporting Information), we used time-series data of fish abundance at XX sites during a period of 1999-2019.

**Fish stocking.** At the protected watersheds, masu salmon were stocked to enhance the wild populations since XXX. A mixture of captive-bred and wild-born individuals were released in spring (fry [age 0+] and smolt [age 1+] stages) and fall (juvenile stage [age 0+]). We examined historical records of fish stocking and obtained annual stocking data (the number of individuals released) for a period of 1999-2019 at each watershed. Although fish were released at multiple locations within a watershed, we were unable to extract stocking data at the site level.

**Environmental data.** At each sampling site, we collected the following data as potential covariates: upstream watershed area (km^2^; a proxy for stream size), proportional land use in the upstream watershed (forest, agriculture, urban), local climates (annual mean air temperature [unit] and cumulative precipitation [mm]), and ocean productivity (Chlorophyll a [unit]). Annual data of ocean productivity (2003-2019) were extracted from OceanColor (<https://oceancolor.gsfc.nasa.gov/>) and the averaged value within the 30-km radius of each river mouth was calculated.

@@

Our goal is to compare temporal dynamics of fish communities across sites. However, the data are not directly comparable because of observation errors (e.g., different observers) and missing observations. To confront this challenge, we developed Bayesian state space models for three species groups separately: (1) whole community, summed abundance of all species, (2) enhanced species, abundance of masu salmon, (3) unenhanced species, summed abundance of all species except masu salmon. Bayesian state space models is best suited for our analysis because it can account for observation errors while imputing missing values given the long-term trend at each site. The model is composed of observation and state models as described below.

@@

$n_{s,t}$ is the "true" fish density excluding fish released in the spring, $Fry_{w(s),t}$ the number of salmon fry released in spring in watershed $w$ within which site $s$ is located, and $\beta_s$ the site-specific effect of released salmon fry on the observed fish density. The parameter $e^{\epsilon_{s,t}^{obs}}$ is a random error variable drawn from a log-normal distribution as $\epsilon_{s,t}^{obs} \sim Normal(0, \sigma_{obs,s}^2)$. The inclusion of this random variable allows us to account for site-specific and year-specific observation errors. The parameter $\psi$ equals zero when we model the unenhanced species group (otherwise $\psi = 1$). We included $\psi$ to exclude the term $\beta_s~Fry_{w(j),t}$ because the unenhanced species group does not include masu salmon.

@@

$$
ln~n_{s,t+1} = ln~r_{s} + ln~n_{s,t} + \epsilon_{s,t}^{state}
$$where $ln~r_{s}$ is the site-specific rate of change at site $s$ and $\epsilon_{s,t}^{state}$ is the process error that is assumed to follow a normal distribution as $\epsilon_{s,t}^{state} \sim Normal(0, \sigma_{state,s}^2)$. The site-specific rate of change is random draws from a normal distribution as $ln~r_s \sim Normal(\mu_r, \sigma_r^2)$, assuming that community dynamics across Hokkaido have a shared temporal trend to some degree (the degree of shared trend is controlled by the standard deviation $\sigma_r$). This hierarchical structure allows for improved estimates of $ln~r_s$ by sharing information across sites. We used median estimates of fish density $n_{s,t}$ to calculate temporal CV, mean ($\mu$), and SD ($\sigma$).
